\documentclass{beamer}
%Information to be included in the title page:
\title{Stability, interpretability and nonprojectibility}
\author{Jayde Massmann}
\date{\today}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}

\usepackage{tikz}
\usetheme{Copenhagen}
\usecolortheme{default}

\usepackage{parskip}

\expandafter\def\expandafter\insertshorttitle\expandafter{%
  \insertshorttitle\hfill%
  \insertframenumber\,/\,\inserttotalframenumber}

\newcommand{\Ord}{\mathsf{Ord}}
\newcommand{\ZFC}{\mathsf{ZFC}}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}
\frametitle{Reflection principles}
Reflection principles and inaccessible cardinals turn out to be a major driving theme in modern set theory. \pause

\begin{itemize}
    \item The universe of all sets is stratified via the von Neumann hierarchy $\langle V_\alpha: \alpha \in \Ord \rangle$.
    \item Lévy-Montague reflection: even without having to assume the existence of inaccessible cardinals etc, $\Ord$ has powerful reflection properties.
    \item For any first-order formula, any set is contained in some object which ``reflects'' that formula, i.e. satisfies it iff the whole universe does.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Formal reflection}
If $\kappa$ is an inaccessible cardinal, then $V_\kappa$ satisfies all the axioms of ordinary set theory, $\ZFC$, and so this reflection principle holds inside $V_\kappa$. We can actually do better, by permitting special predicate symbols (i.e. extending the language of set theory by a ``black box'' oracle).

\begin{block}{Definition}
Let $n, m < \omega$. Then a cardinal $\kappa$ is called $\Pi^n_m$-indescribable if, whenever $\varphi$ is a $\Pi_m(\dot{A})$ sentence and $A \subseteq V_\kappa$ is so that $\langle V_{\kappa+n}, \in, A \rangle \models \varphi$, there is $\alpha < \kappa$ so that $\langle V_{\alpha+n}, \in, A \cap V_\alpha \rangle \models \varphi$.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Notes}
$n$ is meant to be the ``order'', as e.g. a first-order sentence over $V_{\kappa+1}$ becomes a second-order sentence over $V_\kappa$. So the following are basically equivalent: $\kappa$ is $\Pi^n_0$-indescribable, and the $n$-th order reflection principle with a predicate holds in $V_\kappa$. In particular:

\begin{block}{Lemma}
$\kappa$ is inaccessible iff it is $\Pi^1_0$-indescribable.
\end{block}

However, for all $n < \omega$, the least $\Pi^1_n$-indescribable cardinal is not $\Pi^1_{n+1}$-indescribable.
\end{frame}

\section{Admissibility}

\begin{frame}
\frametitle{Admissible sets}
\begin{itemize}
    \item At the countable level, analogues to regular or indescribable cardinals can be obtained by stratifying our universe via the constructible hierarchy instead.
    \item Say $M$ is admissible if all axioms of Kripke-Platek set theory hold within $M$.
    \item For example, $V_{\omega+\omega}$ satisfies all axioms of Kripke-Platek set theory except $\Delta_0$-collection.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Admissible ordinals}
\begin{itemize}
    \item In general, if $V_\kappa$ is admissible, then $\beth_\kappa = \kappa$.
    \item Meanwhile, by using a ``predicative'' construction, and in particular not adding all possible subsets at each next stage, we'll be immune to this counterexample.
    \item The smallest admissible ordinal is equal to $\omega_1^{\mathrm{CK}}$. This is the supremum of order-types of computable well-orders of $\omega$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Generalized recursion theory}
Say $A \subseteq L_\alpha$ is $\alpha$-recursive (resp. $\alpha$-recursively enumerable) if it is $\Delta_1(L_\alpha)$ (resp. $\Sigma_1(L_\alpha)$). The reason for the naming ``$\alpha$-recursive'' is that a subset of $\omega$ is recursive (i.e. computable) iff it is $\Delta^0_1$ in the arithmetical hierarchy, and recursively enumerable iff it is $\Sigma^0_1$.
\end{frame}

\begin{frame}
\frametitle{Recursive analogues}
Admissible ordinals are traditionally considered as a ``recursive analogue'' of regular cardinals because $\alpha$ is admissible iff $\alpha > \omega$, $\alpha$ is a limit ordinal and either of the following hold:

\begin{itemize}
    \item For all $\delta < \alpha$, there is no $\alpha$-recursively enumerable map $\delta \to \alpha$ with cofinal range.
    \item For all $\delta < \alpha$, there is no $\alpha$-recursive surjection $\delta \to \alpha$.
\end{itemize}

This suggests attempting to generalize other large cardinal axioms to the countable level. $\Pi^1_n$-indescribability in particular can be copied almost verbatim, although one has to replace the predicate $A$ with a parameter $b \in L_\alpha$.
\end{frame}

\begin{frame}
\frametitle{Reflecting ordinals}
\begin{block}{Definition}
An ordinal $\alpha$ is $\Pi_n$-reflecting if $\alpha > 0$ and, whenever $\varphi$ is a $\Pi_n$-formula and $b \in L_\alpha$ is so that $L_\alpha \models \varphi(b)$, there is $\beta < \alpha$ so that $b \in L_\beta$ and $L_\beta \models \varphi(b)$.
\end{block}

It's immediate to see that any $\Pi_n$-reflecting ordinal is a limit ordinal. Due to downwards absoluteness, $\Pi_1$-reflection is actually equivalent to being a limit.
\end{frame}

\begin{frame}
\frametitle{$n > 1$}
\begin{itemize}
    \item Meanwhile, $\Pi_2$-reflection is equivalent to admissibility.
    \item For $n > 2$, any $\Pi_n$-reflecting ordinal is a limit of $\Pi_2$-reflecting ordinals.
    \item In general, for $n > 0$, the recursive analogue of $\Pi^1_n$-indescribability may be considered to be $\Pi_{n+2}$-reflection.
\end{itemize}
\end{frame}

\section{Shrewdness and stability}

\begin{frame}
\frametitle{Transfinite extensions}
\begin{itemize}
    \item My work primarily focuses on transfinite extensions of these, i.e. notions beyond $\Pi^n_m$-indescribability or $\Pi_n$-reflection for $n, m < \omega$.
    \item Unfortunately, directly generalizing indescribability or reflection (e.g. simply considering $\Pi^\alpha_m$-indescribability for arbitrary ordinals $\alpha$, without changing the definition) doesn't work.
    \item Instead, we need to slightly modify the definitions. The three notions given in the following definition are, among nonprojectibility and subtlety, going to be the key ones.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Shrewdness, transfinite reflection and stability}
\begin{block}{Definition}
Let $\eta > 0$. A cardinal $\kappa$ is called $\eta$-shrewd iff, for all $P \subseteq V_\kappa$ and every formula $\varphi(x)$ (possibly using the new predicate $\dot{A}$), if $\langle V_{\kappa+\eta}, \in, P \rangle \models \varphi(\kappa)$, then there exist $0 < \kappa_0, \eta_0 < \kappa$ so that $\langle V_{\kappa_0+\eta_0}, \in, P \cap V_{\kappa_0} \rangle \models \varphi(\kappa_0)$.

An ordinal $\alpha$ is called $\xi$-$\Pi_n$-reflecting iff, for all $b \in L_\alpha$ and every $\Pi_n$-formula $\varphi(x)$, if $L_{\alpha+\xi} \models \varphi(b)$, then there exist $\alpha_0, \xi_0 < \alpha$ so that $b \in L_{\alpha_0}$ and $L_{\alpha_0+\xi_0} \models \varphi(b)$.

Let $\xi > 0$. An ordinal $\alpha$ is called $\xi$-stable iff, for all $b \in L_\alpha$ and every $\Sigma_1$-formula $\varphi(x)$, $L_\alpha \models \varphi(b)$ iff $L_{\alpha+\xi} \models \varphi(b)$.
\end{block}
\end{frame}

\section{Introducing and eliminating oracles}

\begin{frame}
\frametitle{$\mathcal{A}$-stability}
Naturally, one may consider adding back an oracle.

\begin{block}{Definition}
Let $\mathcal{A}$ be an arbitrary class and let $\xi > 0$. An ordinal $\alpha$ is called $\mathcal{A}$-$\xi$-stable iff, for all $b \in L_\alpha$ and every $\Sigma_1(\dot{A})$-formula $\varphi(x)$, $\langle L_\alpha, \in, \mathcal{A} \cap L_\alpha \rangle \models \varphi(b)$ iff $\langle L_{\alpha+\xi}, \in, \mathcal{A} \cap L_{\alpha+\xi} \rangle \models \varphi(b)$.
\end{block}

\pause

This notion is trivial when $\mathcal{A} \subseteq L_\alpha$, and obviously most interesting when $\mathcal{A} \subseteq L_{\alpha+\xi}$.
\end{frame}

\begin{frame}
\frametitle{An example}
This notion can attain exorbitant strength -- for example, let $\xi > \alpha$, $L_\xi \models \ZFC$ and $\mathcal{A}$ be the set of $\beta < \xi$ which are $\xi$-stable. \pause Then $\alpha$ is $\mathcal{A}$-$\xi$-stable iff the ordinary definition of $\xi$-stability holds for $\alpha$, but where $\varphi$ is allowed to $\Sigma_2$. That is, $\Sigma_1(\dot{A})$-substructurehood in this case paves the way for full $\Sigma_2$-substructurehood. $\Sigma_2$-stability is actually possibly more relevant to our work than $\Sigma_1$-stability.
\end{frame}

\begin{frame}
\frametitle{What are subtle cardinals?}
Subtle cardinals were introduced by Ronald Jensen in his analysis of the fine structure of $L$.

\begin{block}{Definition}
A cardinal $\kappa$ is subtle if the following holds. Let $C \subseteq \kappa$ be an arbitrary club (closed and unbounded) set. And assume $\vec{S}: C \to \mathcal{P}(\kappa)$ is regressive, in that $\vec{S}(\alpha) \subseteq \alpha$ for all $\alpha \in C$. Then there are $\beta, \delta \in C$ so that $\beta < \delta$ and $\vec{S}(\delta) \cap \beta = \vec{S}(\beta)$.
\end{block}

If $\kappa$ is subtle, then $\lozenge_\kappa$ holds.
\end{frame}

\begin{frame}
\frametitle{Why did we mention them?}
\begin{itemize}
    \item It's known that subtle cardinals are ``small'' large cardinals.
    \item They're below Ramsey, measurable, etc. cardinals (and other ultrafilter- or elementary embedding-based large cardinals) in terms of consistency strength.
    \item But they're above $\Pi^n_m$-indescribable cardinals for all $n, m < \omega$.
    \item So it's natural to try to also calibrate a recursive analogue of them.
    \item It turns out: there's a characterisation of subtlety via shrewdness!
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Extra oracles for shrewdness...}
As one may expect, this characterisation involves adding an extra oracle for shrewdness. 

\begin{definition}
Let $\mathcal{A}$ be an arbitrary class and let $\eta > 0$. A cardinal $\kappa$ is called $\mathcal{A}$-$\eta$-shrewd iff, for all $P \subseteq V_\kappa$ and every formula $\varphi(x)$ (possibly using new predicates $\dot{P}$ and $\dot{A}$), if $\langle V_{\kappa+\eta}, \in, P, \mathcal{A} \cap V_{\kappa+\eta} \rangle \models \varphi(\kappa)$, then there exist $0 < \kappa_0, \eta_0 < \kappa$ so that $\langle V_{\kappa_0+\eta_0}, \in, P \cap V_{\kappa_0}, \mathcal{A} \cap V_{\kappa_0+\eta_0} \rangle \models \varphi(\kappa_0)$.
\end{definition}
\end{frame}

\begin{frame}
\frametitle{Extra oracles for shrewdness...}
We show later that stability with oracles may be considered as a recursive analogue of this version of shrewdness with oracles. For now, let us first give the promised characterisation of subtlety.

\begin{block}{Theorem}
A strongly inaccessible cardinal $\pi$ is subtle iff the following holds: for any $\mathcal{A} \subseteq V_\pi$, there are stationarily many $\kappa < \pi$ so that $\kappa$ is $\mathcal{A}$-$\eta$-shrewd for all $\eta < \pi$.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Recursive subtlety}
This motivates the following definition:

\begin{block}{Definition}
$\rho$ is recursively subtle iff, for any $\rho$-recursively enumerable $\mathcal{A} \subseteq L_\rho$, $\rho$ is $\Pi_2$-reflecting onto the set of $\kappa < \rho$ which are $\mathcal{A}$-$\rho$-stable.
\end{block}

Our main theorem is that recursively subtle ordinals are precisely $\Sigma_2$-nonprojectible ordinals.
\end{frame}

\begin{frame}
\frametitle{Nonprojectibility}
\begin{block}{Theorem}
Suppose $\beta < \omega_1^L$. Then the following are equivalent:

\begin{itemize}
    \item For all $\tau < \beta$, there is $\alpha < \beta$ so that $\tau < \alpha$ and $\alpha$ is $\beta$-$\Sigma_2$-stable.
    \item The $\Sigma_2$-projectum $\rho_2^\beta$ is equal to $\beta$.
    \item If $\mathcal{A} \subseteq \beta$ is $\Sigma_2(L_\beta)$ and $\sup(\mathcal{A}) < \beta$, then in fact $\mathcal{A} \in L_\beta$.
\end{itemize}
\end{block}

\pause

Any ordinal satisfying any of these equivalent conditions is called $\Sigma_2$-nonprojectible (due to condition 2), although only formulation 1 will be relevant for our purposes.
\end{frame}

\begin{frame}
\frametitle{The crux of the proof}
Our goal is so-called oracle elimination: the title of this section. Basically, we want to show that, in $\mathcal{A}$-$\xi$-stability, the additional oracle $\mathcal{A}$ is irrelevant given higher degrees of correctness: any $\xi$-$\Sigma_2$-stable ordinal is $\mathcal{A}$-$\xi$-stable. \pause We don't want to do this for all $\mathcal{A}$, only the $\xi$-recursively enumerable $\mathcal{A}$ (and a bit beyond, including the $\xi$-co-recursively enumerable sets). Without stringent conditions, full oracle elimination is certainly not possible.
\end{frame}

\begin{frame}
\frametitle{Interpretability}
\begin{block}{Definition}
Let $\kappa, \rho$ be ordinals with $\kappa < \rho$. Then $\mathcal{A} \subseteq L_\kappa$ is called $(\kappa, \rho)$-interpretable iff, for any $\Sigma_2$ formula $\varphi$ and parameters $\vec{b} \in L_\kappa$ with $\mathcal{A} = \{x \in L_\kappa: L_\kappa \models \varphi(x, \vec{b})\}$, we have $\mathcal{A} = \mathcal{A}^{L_\rho} \cap L_\kappa$, where $\mathcal{A}^{L_\rho} = \{x \in L_\rho: L_\rho \models \varphi(x, \vec{b})\}$.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Global interpretability from stability}
In other words, $\mathcal{A}$ is $(\kappa, \rho)$-interpretable iff it any $\Sigma_2$-definition gives a way of extending $\mathcal{A}$ to a ``fuller'' subset of $L_\rho$ without adding new elements of $L_\kappa$.

\begin{block}{Lemma}
Let $\kappa, \xi$ be ordinals so that $\xi > 0$. Then the following are equivalent:

\begin{itemize}
    \item $\kappa$ is $\xi$-$\Sigma_2$-stable.
    \item Every subset of $L_\kappa$ is $(\kappa, \kappa+\xi)$-interpretable.
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
\frametitle{Oracle elimination}
Using interpretability, you can derive $\Delta_2(\dot{A})$-preservation from $\Sigma_2$-preservation, i.e. oracle elimination:

\begin{block}{Theorem}
Let $\kappa, \xi$ be ordinals so that $\xi > 0$. Assume $\kappa$ is $\xi$-$\Sigma_2$-stable and $\mathcal{A} \subseteq L_{\kappa+\xi}$ is $\Delta_2$-definable in $L_{\kappa+\xi}$ with parameters from $L_\kappa$. Then $\kappa$ is $\mathcal{A}$-$\xi$-stable.
\end{block}

\end{frame}

\begin{frame}
\frametitle{How to proceed}
Anyways, now we have the following: assume $\rho$ is $\Sigma_2$-nonprojectible. Recall characterisation 1 -- for all $\tau < \rho$, there is $\alpha < \rho$ so that $\tau < \alpha$ and $\alpha$ is $\rho$-$\Sigma_2$-stable. A relatively easy (yet surprising) reflection argument shows that, not only is $\beta$ a limit of $\beta$-$\Sigma_2$-stable ordinals, but actually $\beta$ is $\Pi_2$-reflecting on them. Therefore, for the conclusion that $\rho$ is  recursively subtle, it suffices to prove that a tail of $\rho$-stable ordinals are also $\mathcal{A}$-$\rho$-stable. And to do this, simply pick a $\rho$-stable ordinal $\kappa$ so that $\mathcal{A}$ is definable in $L_\rho$ with parameters from $L_\kappa$.
\end{frame}

\section{Measurable cardinals?}

\begin{frame}
\frametitle{Recursive measurability?}
$\alpha$ is $\Sigma_2$-extendible iff there is $\beta > \alpha$ so that $\alpha$ is $\beta$-$\Sigma_2$-stable.

The smallest $\Sigma_2$-extendible ordinal has a characterisation in terms of infinite-time computability, and arithmetical quasi-inductiveness. It was previously also believed that $\Sigma_2$-extendibility may serve as a recursive analogue of measurability.
\end{frame}

\begin{frame}
\frametitle{Evidence?}
Say $\alpha$ is recursively measurable iff there is a nonprincipal ultrafilter on the Boolean algebra of $\alpha$-recursive subsets of $\alpha$, which is closed under intersections of $< \alpha$ many sets, as long as this intersection can be coded in an $\alpha$-recursive way. Then $\alpha$ is recursively measurable iff it is $\Sigma_2$-extendible. \pause This assignment is however not ``consistent'' with our findings: any measurable cardinal is weakly compact and subtle, but:

\begin{block}{Proposition}
If $\alpha$ is $\Sigma_2$-extendible, it is $\Pi_3$-reflecting and nonprojectible, yet the least $\Sigma_2$-extendible ordinal is not $\Sigma_2$-nonprojectible.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Remark}
Maybe this dissimilarity occurs because there can be no true recursive analogue of measurability. After all, a main application of recursive analogues is in ordinal analysis, which can be carried out absolutely and hence measurable cardinals will never be needed for it.
\end{frame}

\begin{frame}
\frametitle{Thanks!}
Thanks for listening!
\end{frame}

\end{document}
